Going from coarse landings data to fine scale species distribution

### 2022-05-05 Version_M2_plusieurszones/draft/paper_3/paper_3.Rmd
WIP

## An explicitly spatialized catch model

@@

\begin{equation}
\label{eq:defY}
Y_i | S(x_i)  \overset{ind}{\sim} \M{p_i}{\frac{S(x_i)}{1-p_i}}{\sigma^2}
\end{equation}

\begin{align}
Y_i | S(x_i) & \overset{ind}{\sim} p_i \delta_0+ (1-p_i) B_i, \quad B_i \sim \mathcal{L}\left ( \frac{S(x_i)}{1-p_i} ,\sigma^2\right), \cr \nonumber
\mbox{with } p_i &= \exp(-e^\xi S(x_{i}))
\end{align}

@@

To circumvent such limitations, we propose an alternative approach that models the catch declarations $D_j$ instead of the reconstructed individual catch $\Yr(x_i)$. As we  aim to propose a model compatible with the original one, we are lead to specify  the distribution of $D_j$ which consists of a sum of $m_j$ random variables following a mixture distribution. This distribution has no  known analytical form. However the catch declaration data also exhibit some zero inflation and a long tail repartition of the values: a mixture model would be a good candidate. Therefore, we propose to use the same family of distribution for  $D_j$ than the one use for $\Yr$: a mixture between a Dirac mass in $0$ and a lognormal distribution. This is not exact, as the distribution of $D_j$ should be defined as convolution of mixtures but would result in quite untractable model. We somehow approximate the sum of  the original mixture distributions  by a mixture distribution of the same kind and we define the different key quantities, i.e  the mixture proportion, the expected positive catch declaration and its variance so that to match the original propostion:

@@

$$D_j | \mathcal{P}_j,\bm{S} \sim \mathcal{L}_D( \mu_j,\xi,\sigma^2)$$

with $\mathcal{P}_j=(x_1,...,x_i, ...,x_{m_j})$, the list of the fishing points associated to declaration $D_j$,  $\mu_j$ the expected positive biomass, $\xi$ the zero-inflation parameter and $\sigma^2$ the variance parameter.

@@

**Matching $\mathcal{L}_Y$ and $\mathcal{L}_D$**

First, as at the punctual level ($Y$), the declarations can be decomposed in 2 components: (1) the probability to obtain a zero-value declaration and, if the declaration is positive, (2) the probability to obtain some declaration value $d_j$.

(1) We derive the probability to obtain a zero-declaration $P(D_j = 0)$ by simply multiplying the probability to obtain a zero-punctual observation $P(Y_i=0 )$ to all fishing points $x_i \in \mathcal{P}_j$.

\begin{align*}
P(D_j = 0) & = \prod_{i\in \mathcal{P}_j} P(Y_{i} = 0),\nonumber \\
                      & = \exp{ \left \lbrace- \sum_{i\in \mathcal{P}_j} e^{\xi}. S(x_{i})\right \rbrace} = \pi_j.
\end{align*}

(2) If the declaration is positive, we assume the probability to obtain catch $d_j$ follows a continous distribution $\mathcal{L}_{D_j|D_j>0}$ with mean component $\mu_j$ and variance component $\sigma_j$. From $\mathcal{L}_Y$ we can derive $E(D_j \vert D_j > 0)$ and $Var(D_j \vert D_j > 0)$ (given below) and re-express $\mu_j$ and $\sigma_j$ as a function of these two moments.

$$E(D_j \vert D_j > 0)=\frac{\sum_{i \in \mathcal{P}_j} S(x_{i})}{1-\pi_j}$$
$$Var(D_j \vert D_j > 0) = \frac{\sum_{i \in \mathcal{P}_j} Var(Y_{i})}{1-\pi_j} - \frac{\pi_j}{(1-\pi_j)^2}E(D_j)^2$$
$$Var(Y_{i})=\frac{S(x_{i})^2}{1-p_{i}}(e^{\sigma^2}-(1-p_{i}))$$

If we consider $\mathcal{L}_{D_j|D_j>0}$ is lognormal (with parameterization described in SM) then, $\mu_j = E(D|D_j>0)$ and $\sigma_j = ln(\frac{Var(D|D_j>0)}{E(D|D_j>0)^2} + 1)$. Such way, by fitting the model on declaration $D_j$ rather than reallocated catch $Y_i$ and linking $\mathcal{L}_D$ to $\mathcal{L}_Y$, we intend to better follow the observation process of the commercial data and potentially improve inference compared with the approximation realized through uniform reallocation. All calculations to obtain these formulas are available in SM.

@@

@@

**Single-square simulations**

@@

- a configuration corresponding to the actual situation (uniform reallocation): simulated catches are summed into declarations and reallocated over the related fishing locations. We then fit the $Y_i^*$ model (the model fitted to punctual reallocated data).

@@

**Expectancy of a positive declaration**

>>>>

## A spatialized catch model for aggregated data

@@

\begin{equation}
\label{eq:defY}
Y_i | S(x_i)  \overset{ind}{\sim} \M{p_i}{\mu_i}{\sigma^2}
\end{equation}
with $p_i \defeq \exp(-e^\xi S(x_{i})$ the proportion of the mixture, $\mu_i\defeq 
\frac{S(x_i)}{1-p_i}$ the expected catch when positive and $\sigma^2$ its variance (this parametrization of a lognormal distribution is detailed in SMXXXX and corresponds to

\begin{align}
\P\left (Y_i =0 | S(x_i) \right) & = p_i=  \exp(-e^\xi S(x_{i}) ) \\
\E\left ( Y_i \vert Y_i >0 , S(x_i)\right)  & =  \frac{S(x_i)}{1-p_i},\\  \nonumber
\V ar( \log{Y_i}\vert Y_i >0, S(x_i) )& = \mu_i^2 (e^{\sigma^2}-1), \cr \nonumber
\end{align}

@@

To circumvent such limitations, we propose an alternative approach that models the catch declarations $D_j$ instead of the reconstructed individual catch $\Yr(x_i)$. As we  aim to propose a model compatible with the original one, we are lead to specify  the distribution of $D_j$ which consists of a sum of $m_j$ random variables following a mixture distribution. This distribution has no  known analytical form. However the catch declaration data also exhibit some zero inflation and a long tail repartition of the values: a mixture model would be a good candidate. Therefore, we propose to use the same family of distribution for  $D_j$ than the one use for $\Yr$: a mixture between a Dirac mass in $0$ and a lognormal distribution. This is not exact, as the distribution of $D_j$ should be defined as convolution of mixtures which  would result in an untractable model. We somehow approximate the sum of  the original mixture distributions  by a mixture distribution of the same kind and we define the different key quantities, i.e  the mixture proportion, the expected positive catch declaration and its variance so that to match the original propostion:

@@

$$D_j | \mathcal{P}_j,\bm{S} \sim \mathcal{L}_D( p^D_j, \mu^D_j, \sigma^{2,D})$$

with $\mathcal{P}_j=(x_1,...,x_i, ...,x_{m_j})$, the list of the fishing points associated to declaration $D_j$,  $\mu^D_j$ the expected positive biomass, $p_j^D$ the proportion of the mixture and $\sigma^{2,D}$ the variance parameter.

@@

**Matching the individual catch and the declaration levels**

In order to define compatible models for $Y$ the actual individual catches and $D$ the catch declarations, we choose to match the key quantities of the two distributions.

(1) As the $Y_1, \ldots, Y_{m_j}$ are independent conditionally on $\bm{S}$,  the probability to obtain a zero-declaration $P(D_j = 0)$ is obtained by simply multiplying the probability to obtain a zero-punctual observation $P(Y_i=0 )$ to all fishing points $x_i \in \mathcal{P}_j$.

\begin{align*}
\P(D_j = 0\vert \bm{S}, x_1, \ldots x_{m_j})  & = \prod_{i\in \mathcal{P}_j} P(Y_{i} = 0\vert \bm{S}, x_i),\nonumber \\
                      & = \exp{ \left \lbrace- \sum_{i\in \mathcal{P}_j} e^{\xi}. S(x_{i})\right \rbrace} = \pi_j.
\end{align*}

(2) The continuous component of the mixture is defined by the expected mean of a positive declaration and a transformation of its variance. It is straightforward to prove (see  SMXXXXX) that

$$\E(D_j \vert D_j > 0)= \frac{\sum_{i \in \mathcal{P}_j} S(x_{i})}{1-\pi_j}$$
$$\V ar(D_j \vert D_j > 0) = \frac{\sum_{i \in \mathcal{P}_j} \V ar(Y_{i})}{1-\pi_j} - \frac{\pi_j}{(1-\pi_j)^2}E(D_j)^2$$
$$\V ar(Y_{i})=\frac{S(x_{i})^2}{1-p_{i}}(e^{\sigma^2}-(1-p_{i}))$$

By using this second model, we expect to represent more accurately the observation process and avoid some of the drawbacks of the estimationbased on reallocated indivdual catch data $Y^r$.

@@

XXXX  Il faut aussi présenter la partie commercial ou dire que c'est comme dans Alglave 2022

@@

**Single-square simulations**
XXX Infor sur le nombre de simus ???

@@

- a configuration corresponding to the actual situation (uniform reallocation): simulated catches are summed into declarations and reallocated over the related fishing locations. We then fit the $Y_i^r$ model (the model fitted to individual reallocated catch data).

@@

**Expectation of a positive declaration**
